Segmentation-registration of diffusion MR images of the brain driven by active-contours

### 2015-02-17 Publications/2015-NeuroImage/04-methods.tex
write phantom experiments. close #222

@@

\paragraph*{Simulated phantoms}%
\label{sec:digital_phantoms}
We proved the concept on simplistic digital phantoms which we will name after their
  appearance as ``box'', ``ball'', ``L-shape'', and ``gyrus'' (\autoref{fig:phantom}).
Using \emph{phantomas} \citep{caruyer_phantomas_2014}, we simulated \gls*{t1} and
  \gls*{t2} images (TE/TR=10/1500ms and TE/TR=90/5000ms
  respectively) corresponding to each shape.
Simulations were conducted with rician noise for a \gls*{snr} of 300.0, and
  two image resolutions, $1.0mm$ and $2.0mm$, both isotropic.
The reference surfaces were extracted using the marching-cubes algorithm
  shipped with \emph{FreeSurfer} \citep{fischl_freesurfer_2012}
  at the higher resolution (1.0mm isotropic)..

\begin{figure}
	\includegraphics[width=\linewidth]{figures/figure02}
	\caption{The ``cortex'' phantom is a spherical shape with two sulci and an
	 outer crust resembling the cortical folding.
	When the model is used to generate low resolution images, the
	  partial volume effect turns segmentation of the sulci a
	  challenging problem with voxel-wise clustering methods,
	  but it is successfully segmented with the proposed method (see
	  \autoref{fig:results_phantom}).
	Top row shows the central-axial plane of the simulated images at
	  high resolution and noiseless (left is \gls*{t1} and right is
	  \gls*{t2}).
	Bottom row represents the extracted contours that are used as shape
	  priors to be registered to low-resolution, noise corrupted versions
	  of the phantom.
	}\label{fig:phantom}
\end{figure}

>>>>

{\color{red}
\paragraph*{Convergence evidence and settings}
 discuss choice of $\tau$, Courant-Friedrichs-Lewy (CFL) condition / Wolfe conditions, multi-resolutions on the bspline grid, image samples and surface sampling,  etc.
\paragraph*{Efficient field densification}
As long as the dense deformation field is iteratively interpolated
from the same set of control points that define the $L-1$ contours,
it is possible to pre-cache all the $\psi_k(\vec{v}_i)$ weights into a
sparse matrix for fast densification. As well, we achieve a
diffeomorphic transform by}

@@

\paragraph*{Simulated phantoms}%
\label{sec:digital_phantoms}
We proved the concept on simplistic digital phantoms which we will name after their
  appearance as ``box'', ``ball'', ``L-shape'', and ``gyrus'' (\autoref{fig:phantom},
  boxes A and B).
Using \emph{phantomas} \citep{caruyer_phantomas_2014}, we simulated \gls*{t1} and
  \gls*{t2} images (TE/TR=10/1500ms and TE/TR=90/5000ms
  respectively) corresponding to each shape.
Simulations were conducted with rician noise for a \gls*{snr} of 300.0, and
  two image resolutions, $1.0mm$ and $2.0mm$, both isotropic.
The reference surfaces were extracted using the marching-cubes algorithm
  shipped with \emph{FreeSurfer} \citep{fischl_freesurfer_2012}
  at the higher resolution (1.0mm isotropic).

\begin{figure*}
	\includegraphics[width=\linewidth]{figures/figure02}
	\caption{A. The ``cortex'' phantom is a spherical shape with two sulci and an
	  outer crust resembling the cortical folding (left).
	The model is used to generate \gls*{t1} and \gls*{t2} images after warping the
	  contours using a random and plausible transformation $U_{true}^{-1}$ (right).
	B. Visual assessment of the results on the low resolution sets:
	  ``gyrus'' (top-left), ``L-shape'' (top-right), ``ball'' (bottom-left),
	  and ``box'' at (bottom-right).
	In yellow color, the recovered contours after registration are represented.
	Our method showed high accuracy, as they are overlapping the ground truth surfaces
	  depicted in green.
	Partial volume effect turns segmentation of the sulci a challenging problem with voxel-wise
	  clustering methods, but it is successfully segmented with our method.
	C. Quantitative evaluation of registration error in terms of average Hausdorff distance of
	  surfaces at high (left) and low (right) resolutions, demonstrating that the error is
	  consistently below the image resolution.
	  }\label{fig:phantom}
\end{figure*}

### 2015-02-17 Publications/2015-NeuroImage/06-results.tex
write phantom experiments. close #222

\subsection*{Proof of concept}
\label{sec:results_phantom}

Applying the evaluation protocol introduced in \autoref{sec:experiments_evaluation} on
  the ``cortex'' phantom (\autoref{fig:phantom}), we successfully tested the method and
  fine tuned the algorithm for its use on real data.
In this characterization experiment, we generated random smooth displacement fields $U^{-1}_{true}$
  (step 4 in \autoref{sec:experiments_evaluation}) based on a B-Spline field with
  control points evenly located in a $6 \times 6 \times 6$ grid covering the region of the phantom.
Invertibility is ensured by controlling the maximum	displacement
  \citep{rueckert_diffeomorphic_2006}.
The software tool to generate these fields is included with the released package.
In this experiment we used $\hat{U}_{cc} = (U^{-1}_{true})^{-1}$, i.e. the ground-truth
  displacement field.
We performed the experiment in two different realizations of the phantom, one at full resolution
  and a second with half resolution to study the performance under strong partial volume effect
  conditions.
The distorted phantoms and visual results are presented in \autoref{fig:results_phantom}.
Quantitative evaluation of the \gls*{swindex} is reported in \autoref{tab:swi_phantom}.

\begin{table}
	\caption{Quantitative evaluation on phantoms: \acrlong*{swindex}}
	\label{tab:swi_phantom}
    \begin{tabular}{llll}
    Voxel size  & Interior surface & Exterior surface & Average \\
    $1.0mm$ & $0.718mm$            & $0.724mm$            & $0.721mm$   \\
    $2.0mm$ & $0.723mm$            & $0.726mm$            & $0.725mm$   \\
    \end{tabular}
\end{table}

{\color{red}
\paragraph*{Convergence evindence and settings}
 discuss choice of $\tau$, Courant-Friedrichs-Lewy (CFL) condition / Wolfe conditions, multi-resolutions on the bspline grid, image samples and surface sampling,  etc.
\paragraph*{Efficient field densification}
As long as the dense deformation field is iteratively interpolated
from the same set of control points that define the $L-1$ contours,
it is possible to pre-cache all the $\psi_k(\vec{v}_i)$ weights into a
sparse matrix for fast densification. As well, we achieve a
diffeomorphic transform by}

% In regular intervals, after certain number of iterations,
% the parameters $\Theta_l$ of the regions can be reestimated
% based on the shifted volumetric samples
% $\vec{x}' = \vec{x_0} + u(\vec{x})$.

\begin{figure*}
\includegraphics[width=\textwidth]{figures/figure03}
\caption{Proof of concept of the registration method. First row shows the registration experiment
  on the high-resolution ($1.0mm$ isometric) realization of the phantom, second row contains the
  low-resolution ($2.0mm$ isometric) version.
  Both phantoms were simulated with \gls*{snr} 30.
  In the first column, the distorted phantoms presented with the undistorted contours.
  For the second column, the contours have been replaced by the distorted contours using
    the ground-truth warping $U^{-1}_{true}$.
  The third column contains again the contours warped with the ground-truth field
  (thin yellow line), and the corresponding contours (green and blue thick lines)
  obtained with the presented method.
  Finally, last row represents the recovered images unwarped using the $\hat{U}_{tst}$
    found through our registration.
  }\label{fig:results_phantom}
\end{figure*}

\subsection*{Susceptibility distortion correction}
\label{sec:results_hcp}

>>>>

\subsection*{Proof of concept on digital phantoms}
\label{sec:results_phantom}
Using the experimental instrumentation described in section \nameref{sec:experiments_evaluation},
  we fine tuned the algorithm for its use on real data and collected quantitative results.
To warp the simulated datasets, we generated 150 realizations of random and smooth displacement
  fields $U^{-1}_{true}$.
Smoothness is achieved using two levels of B-Spline functions, with control points evenly
  located in isotropic grids covering the full extent of the model.
The first level had $50.5mm$ of separation between control points and the second $25.25mm$.
The rationale behind this choice is producing large deformations on the contours with
  the first level, and matching the properties of susceptibility-derived distortions
  as previously reported by \cite{irfanoglu_susceptibility_2011} with the finer grid.
Invertibility of $U^{-1}_{true}$ is ensured by controlling the maximum displacement at each
  level ($20.2mm$ and $10.1mm$ respectively) as studied in \citep{rueckert_diffeomorphic_2006}.

A total of 1200 registration experiments (4 phantom types $\times$ 150 random warpings
  $\times$ 2 resolutions) were run.
For each registration, the averaged Hausdorff distance (see \nameref{sec:experiments_evaluation})
  of the inner and the outer surfaces was computed.
The results showed a consistent and high accuracy, below voxel resolution.
\autoref{fig:phantom} (block C) presents the box plots by model type corresponding
  to the two sets of resolutions of generated phantoms.
To support that the missregistration errors averaged per experiment was significantly
  under the resolution of the target image, we proceeded as follows.
First, we confirmed that the error distributions were skewed using a Shapiro-Wilk test of
  normality.
All the distributions of errors under test (4 phantom types $\times$ 2 resolutions) resulted
  non-normal with $p<0.001$.
Consequently, we used the non-parametric Wilcoxon signed-rank test along with Bonferroni
  correction for multiple comparisons ($N=150$).
Averaged errors resulted significantly lower than voxel resolution with $p < (0.001 / 150)$
  for all the tests (4 phantom types $\times$ 2 resolutions).

\subsection*{Distortion correction on real datasets}
\label{sec:results_hcp}
